Estimation of the road traffic sound levels based on Non-Negative Matrix Factorization

### 2018-01-04 paper/gloaguenNmfTraffic.tex
add readme and NMFthreshold experience

\author{
    Jean-Rémy GLOAGUEN\\
    Arnaud Can\\
    LAE\\
    Ifsttar\\
    jean-remy.gloaguen@ifsttar.fr
  \and
    Mathieu Lagrange\\
	Jean-François Petiot \\
    LS2N, CNRS\\
    Ecole Centrale de  Nantes\\
}
\date{}
\begin{document}

@@

\begin{figure*}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=.45\linewidth]{figures/bloc_diagram_NMF_EN.pdf}
        \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
        \label{fig:TIR_class_filter}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=.45\linewidth]{figures/bloc_diagram_filtrage_EN.pdf}
        \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter \ref{fig:blocDiagramFilter} and for NMF \ref{fig:blocDiagramNMF}}
		\label{fig:TIR_class_sup}
    \end{subfigure}
\end{figure*}

@@

@@

The first estimator to determine the traffic sound level is a frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz. The spectrograms $\mathbf{V}$ are filtered and the remaining energy is then considered as traffic component (eq. \ref{eq:v_tr_filtered}).

@@

The second estimator is the proposed scheme, based on several flavors of the NMF framework. Multiples experimental factors are involved here between the dictionary learning and the own parameters of the NMF.

@@

Table \ref{tab:dictionary_factors} summarizes the parameters and their related values.

\begin{table}[h]
\centering
\begin{tabular}{cccc}
Parameter &  \multicolumn{3}{c}{value}\\ \hline
$K$ & 25  & 50 & 100 \\ \hline
$\mathbf{w_t}$ (s)& 0.5 & 1  & \textit{all}
\end{tabular}
\caption{Summary of the dictionary parameters}
\label{tab:dictionary_factors}
\end{table}

\subsubsection{NMF parameters}

@@

With the third octave spectral representation, only the case $f_c = 20$ kHz is applied. For the TI-NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step. Table \ref{tab:estimation_factors} summarizes the parameters and their related values.

\begin{table*}[]
\centering
\begin{tabular}{lcccccc}
\textbf{Experimental factors} &  \multicolumn{6}{c}{value}\\ \hline
$\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\ \hline
\textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\ \hline
$\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\ \hline
$\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ \hline
\textbf{method} & \multicolumn{2}{c}{filter} & supervised NMF & semi-supervised NMF & TI-NMF \\ \hline
$\mathbf{t}$ & \multicolumn{6}{c}{0.20:0.01:0.70} \\ \hline
\end{tabular}
\caption{Summary of the different parameters taken into account in the frequency low-pass filter and NMF process and their values for the estimation of the traffic sound level}
\label{tab:estimation_factors}
\end{table*}

@@

In all, according Tables \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ (table \ref{tab:dictionary_factors}) and the multiple parameters taken into account by NMF:  for the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for supervised and semi-supervised, 7560 associations of parameters are made. For TI-NMF, 16800 combinations can be calculated.

@@

Table \ref{tab:results} summarized, according to the 3 main parameters (method, $\beta$ and spectral representation), the $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).\\

\ml{replace parameters by experimental factors}

@@

Between the dictionary parameters ($K$ and $w_t$) and the cut-off frequency $f_c$, none is common to all the settings. If for the third octave spectral representation, $f_c$ has been fixed at 20 kHz, in linear representation, this parameters is variable depending on chosen method.

For supervised NMF, the spectrogram $\mathbf{V}$ has to be filtered at 500 Hz to focused the reconstruction of the signals on the part where most of the traffic energy is. According to the choice of $\beta$, the dictionary parameters are different. It is the only case where it is.
In the opposite, in semi-supervised NMF, an unique set of parameters is used to get high performance : $K$ = 100, $w_t$ = 0.5 second and $f_c$ = 20 kHz. It is with $f_c = 20$ kHz that the approximation is better as it includes easily other sound sources. With a lower cut-off frequency ($ f_c <$ 5 kHz), focusing on traffic energy, the risk is then to include \textit{traffic} component in $\mathbf{W_r}$ and to deteriorate the traffic signal reconstruction. The influence of the spectral representation and $\beta$ is here weak as the errors are similar with a low standard deviation. The third approaches, TI-NMF, proposed, in the four cases, the lowest errors. Like supervised NMF, in linear spectral representation, it is necessary to filter $\mathbf{V}$ to focus on the traffic energy. Here $f_c$ is higher (5000 Hz and 2000 Hz respectively for $\beta = 1$ and $\beta$ = 2). The choice of the spectral representation influence also the order of magnitude of the threshold $t$: in the interval [0.37 0.39] and [0.54 0.57] respectively for a linear and third octave spectral representation.\\

>>>>

\author{
    Jean-Rémy GLOAGUEN\\
    Arnaud Can\\
    LAE\\
    Ifsttar\\
    jean-remy.gloaguen@ifsttar.fr
  \and
    Mathieu Lagrange\\
	Jean-François Petiot \\
    LS2N, CNRS\\
    \'Ecole Centrale de  Nantes\\
}
\date{}
\begin{document}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.7\textwidth]{figures/bloc_diagram_source_separation.pdf}
\caption{Bloc diagram of the source separation method}
\end{figure*}

@@

The experience is lead with the Matlab software. In order to be a reproductive experience, all the programs\footnote{\url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}} and the sound database\footnote{\url{https://sandbox.zenodo.org/record/176695#.Wk4ow3kiGos}} (TEMPORAIRE) used are available online.

@@

The first estimator to determine the traffic sound level is a frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz (see Figure \ref{fig:bloc_filtre}). The spectrograms $\mathbf{V}$ are filtered and the remaining energy is then considered as traffic component (eq. \ref{eq:v_tr_filtered}).

@@

The second estimator is the proposed scheme, based on several flavors of the NMF framework. Multiples experimental factors are involved here between the dictionary learning and NMF (see bloc diagram in Figure \ref{fig:bloc_nmf}).

\begin{figure}

    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/bloc_diagram_filtrage_EN.pdf}
  		\caption{}
  		\label{fig:bloc_filtre}
    \end{subfigure}%
    \vfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/bloc_diagram_NMF_EN.pdf}
		\caption{}
		\label{fig:bloc_nmf}
    \end{subfigure}

    \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter (a) and for NMF (b)}
\end{figure}

@@

Table \ref{tab:dictionary_factors} summarizes the experimental factors and their related values.

\begin{table}[h]
\centering
\begin{tabular}{cccc}
Parameter &  \multicolumn{3}{c}{value}\\ \hline
$K$ & 25  & 50 & 100 \\ \hline
$\mathbf{w_t}$ (s)& 0.5 & 1  & \textit{all}
\end{tabular}
\caption{Summary of experimental factors of the dictionary}
\label{tab:dictionary_factors}
\end{table}

\subsubsection{Experimental factors of the NMF}

@@

With the third octave spectral representation, only the case $f_c = 20$ kHz is applied. For the TI-NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step. Table \ref{tab:estimation_factors} summarizes the experimental factors and their related values.

\begin{table*}[]
\centering
\begin{tabular}{lcccccc}
\textbf{Experimental factors} &  \multicolumn{6}{c}{value}\\ \hline
$\mathbf{TIR}$ (dB) & -12 & -6 & 0 & 6 & 12 &  \\ \hline
\textbf{sub-classes} & alert & animals & climate & humans & transportation & mechanics \\ \hline
$\mathbf{\beta}$ & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} \\ \hline
$\mathbf{f_c}$ (kHz) & 0.5 & 1 & 2 & 5 & 10 & 20 \\ \hline
\textbf{method} & \multicolumn{2}{c}{filter} & supervised NMF & semi-supervised NMF & TI-NMF \\ \hline
$\mathbf{t}$ & \multicolumn{6}{c}{0.20:0.01:0.70} \\ \hline
\end{tabular}
\caption{Summary of the different experimental factors taken into account in the frequency low-pass filter and NMF process and their values for the estimation of the traffic sound level}
\label{tab:estimation_factors}
\end{table*}

@@

In all, according Tables \ref{tab:dictionary_factors} and \ref{tab:estimation_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ (table \ref{tab:dictionary_factors}) and the multiple experimental factors taken into account by NMF:  for the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for supervised and semi-supervised, 7560 associations of factors are made. For TI-NMF, 16800 combinations can be calculated.

@@

Table \ref{tab:results} summarized, according to the 3 main factors (method, $\beta$ and spectral representation), the $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all).\\

@@

Between the dictionary factors ($K$ and $w_t$) and the cut-off frequency $f_c$, none is common to all the settings. If for the third octave spectral representation, $f_c$ has been fixed at 20 kHz, in linear representation, this factors is variable depending on chosen method.

For supervised NMF, the spectrogram $\mathbf{V}$ has to be filtered at 500 Hz to focused the reconstruction of the signals on the part where most of the traffic energy is. According to the choice of $\beta$, the dictionary factors are different. It is the only case where it is.
In the opposite, in semi-supervised NMF, an unique set of factors is used to get high performance : $K$ = 100, $w_t$ = 0.5 second and $f_c$ = 20 kHz. It is with $f_c = 20$ kHz that the approximation is better as it includes easily other sound sources. With a lower cut-off frequency ($ f_c <$ 5 kHz), focusing on traffic energy, the risk is then to include \textit{traffic} component in $\mathbf{W_r}$ and to deteriorate the traffic signal reconstruction. The influence of the spectral representation and $\beta$ is here weak as the errors are similar with a low standard deviation. The third approaches, TI-NMF, proposed, in the four cases, the lowest errors. Like supervised NMF, in linear spectral representation, it is necessary to filter $\mathbf{V}$ to focus on the traffic energy. Here $f_c$ is higher (5000 Hz and 2000 Hz respectively for $\beta = 1$ and $\beta$ = 2). The choice of the spectral representation influence also the order of magnitude of the threshold $t$: in the interval [0.37 0.39] and [0.54 0.57] respectively for a linear and third octave spectral representation.\\
